---
title: "Day2Notes"
author: "Wen Zhang"
date: "September 13, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#R Packages

ROpenSci is a good resource to look at for package standards-- quite extensive code review for submitted packages.  

####Common things found in R Packages:  

1. R Code
2. Data
3. Documentation (man); function documentation
4. Namespace
5. Description (metadata)
6. Vignettes; describe how the pieces of package fit together
    - R Code (extracted)
    - HTML
7. Source (compiled) code, e.g. C, C++, Fortran
    - Feasible to learn enough C or C++ within a few weeks to write useful code.
8. Tests
    - Basically double-entry bookkeeping-- describe function and test outputs using ref file(s)/test(s)


```{r findPkgFuns, results="hold", eval=FALSE}
library(ggplot2)
help(package = ggplot2)
#or
ls("package:ggplot2")

vignette() #list vignettes
browseVignettes() #as well

data() #list datasets
```

####Finding package source code

cran package source (tar.gz), because cran only allows open source code.  
R code obviously should be in R dir.  
If a function contains compiled C code, can be found in src dir.  

On your local system, for installed packages, you will find .db files-- for more efficient loading.  

Most modern R packages are hosted on github, so if provided, can also go there to explore source code.  


```{r libPaths, results="hold", eval=FALSE}
.libPaths() #find where packages are installed
```

####Types of package

There are two types of packages:  

1. Packages you install from cran
2. Packages installed in your library (which can be from cran or local or github, etc.).  

Source --> Bundle --> Binary          --> Installed   --> In memory  
files  --> tar.gz --> cached/compiled --> install cmd --> library()  

Today, we will cover `load_all()`.  
Also Build and Reload.  

```{r rv2Install, results="hold", eval=FALSE}
#rv2 is the complete package
#test-me is one of the incomplete versions for learning.
#Just install for now.

#devtools::install("C:/Users/T450S/Desktop/RWorkShop2016/data/day-2/rv2")
#help(package = rv2)

```

####Best practices for package naming:  

1. Only lowercase letters and numbers
2. Maybe add an 'r'
3. Or find a related word and modify
4. Easy to google/search on
    - Weird names are good/specific
5. Memorable

Once you have a name you can create the package.  

```{r createPkg, results = "hold", eval=FALSE}
devtools::create("path/to/package")
devtools::setup("existing/path/")

#Both devtools and RStudio can be used. Both are fine-- we are going to use devtools today. 
#Do not use package.skeleton().

#Create a new package called rv3
#devtools::create("C:/Users/T450S/Desktop/RWorkShop2016/rv3")
#CTRL + . to quickly find function
#Obviously, in rv3, /R is going to be blank upon creation.  
```

####Once you have R code (roxygenated), you can use a shortcut to load and use

Note that you don't need to use `install()` or `document()` to actually load the R code.  
This is quite handy.  

```{r loadAll, results = "hold", eval=FALSE}

#Navigate to the included, complete rv2/R dir
#Copy the R files into your new rv3/R dir
#Open rv3.Rproj

devtools::load_all(".") #Or use KB Shortcut, CTRL + SHIFT + L
rv3 <- dice(1:6)

```


####Function purity

Impure functions = things that create side effects/affect the environment.  
    - `library()`, `print()`, 'install.packages()`, `attach()`
    - All the rng functions as well, since they do not only depend on input
       - Or `Sys.time()`, `read.csv()`  
       
Things to avoid:  

1. Instead of print(), use message(), to let users do things like:  
    - `library(ggplot2, quietly = TRUE)`
    - `suppressMessages(library(data.table))`

```{r msgOptions, eval=FALSE}
if (!quiet) {
message("Processing file ", path)
}
if (is.null(by)) {
by <- ...
message("Joining by ", paste(by, collapse = ", "))
}

```

Impure functions have very important functions. But don't use them in your packages as a rule, unless 
this is what the purpose of your function is. 


####Options

If you must, capture current options at start, reset to original on exit.  

```{r opts, eval=FALSE}
# Bad!
options(stringsAsFactors = FALSE)
read.csv(path)
# Better
old <- options(stringsAsFactors = FALSE)
on.exit(option(old), add = TRUE)
read.csv(path)
# Best: use stringsAsFactors arguments to specific
# functions (not always possible)
read.csv(path, stringsAsFactors = FALSE)

```


```{r onExit, eval=FALSE, results="hold"}
#on.exit() runs regardless of how function exits (even if it fails-- unless R crashes)

f <- function(x) {
  on.exit(message("Hi!"), add = TRUE) #default is FALSE, so get into habit of setting to TRUE
  if (x < 0) {
    stop("!")
  } else {
    10
  }
}
f(-10)
f(10)

#If you have multiple on.exit calls, add = TRUE is particularly important to set. 
```


Clean up after yourself if you change things.  

```{r, eval=FALSE, results="hold"}
f <- function(x) {
  tmp <- tempfile()
  on.exit(unlink(tmp), add = TRUE)
  old_par <- par(bg = "red")
  on.exit(par(old_par), add = TRUE)
  plot(1:10)
}
```


Don't mix side-effects and computation.  

```{r, eval=FALSE, results="hold"}
fortify.lm <- function(model, data = model$model, ...) {
  infl <- influence(model, do.coef = FALSE)
  data$.hat <- infl$hat
  data$.sigma <- infl$sigma
  data$.cooksd <- cooks.distance(model, infl)
  data$.fitted <- predict(model)
  data$.resid <- resid(model)
  data$.stdresid <- rstandard(model, infl)
  data
}

#Also

mod <- lm(mpg ~ wt, data = mtcars)
plot(mod)
getS3method("plot", "lm") #Gets statistics and visualizes them. Quite long. Compare to fortify.lm
#Makes reuse of code very difficult. 
#ggplot2 takes a different approach-- viz is independent of computation. 
```


Improve this function:  
```{r, eval=FALSE, results="hold"}
# plot_function <- function(f, xlim = c(0, 1), n = 100) {
#   x <- seq(xlim[1], xlim[2], length.out = n)
#   y <- f(x)
#   print(paste0("Using ", n, " points"))
#   par(bg = "grey90")
#   plot(x, y, xlab = "x", ylab = "f(x)", type = "l")
# }
# plot_function(sin)
# plot(runif(5))
```

One way to do this, to at least reset color par option, although still munging together data and plot:  
    - And using message instead of print.  


    
```{r, eval=FALSE, results="hide"}
plot_function_2 <- function(f, xlim = c(0, 1), n = 100) {
  
  x <- seq(xlim[1], xlim[2], length.out = n)
  y <- f(x)
  
  
  message("Using ", n, " points") #message does not require paste
  opar <- par("bg") 
  par(bg = "grey90")
  plot(x, y, xlab = "x", ylab = "f(x)", type = "l")
  
  on.exit(par("bg" = opar), add = TRUE)
}

plot_function_2(sin)
plot(runif(5))

suppressMessages(plot_function_2(sin))
plot(runif(5))

```

An alternative, better:  

```{r, eval=FALSE, results="hide"}
grid_function <- function(f, xlim = c(0, 1), n = 100, quiet = FALSE) {
  if (!quiet) {
    message("Using xlim: ", xlim[1], "-", xlim[2])
  }
  if (!quiet) {
    message("Using ", n, " points")
  }
  x <- seq(xlim[1], xlim[2], length.out = n)
  y <- f(x)
  data.frame(x, y)
}

plot_function <- function(f, xlim = c(0, 1), n = NULL) {
  fun <- grid_function(f, xlim, n)
  old <- par(bg = "grey90")
  on.exit(par(old), add = TRUE)
  plot(fun$x, fun$y, xlab = "x", ylab = "f(x)",
       type = "l")
}

```


###DESCRIPTION file

Default template to get you started.  
    - Note that if KB shortcuts do not work in RStudio, it is probably because your package lacks a DESCRIPTION
    
Depends, Imports, and Suggests:  
    - Depends basically to describe what version of R you need (optional version spec)
        - Can use this to also list packages, but this is deprecated and generally should not be used. 
    - Imports are things you NEED-- installed automatically if needed. These are required. 
    - Suggests are nice to have, not installed automatically

Note that packages in Imports are installed, but not necessarily loaded.  
If in DESCRIPTION: Imports: foo  
Then call foo via foo::  

It is probably better to use double colons to be explicit in your package functions.  

If you are using suggested packages:  

```{r, eval=FALSE, results="hide"}
new_function <- function(x, y, z) {
  if (!requireNamespace("foo", quietly = TRUE)) {
    stop("Need foo! Use install.packages('foo').")
  }
  foo::bar(x, y) + z
}
```

Why use Suggests?  
1. Needed for development, e.g. testthat  
2. Heavy dependencies and only needed in special situations  
    * ggplot2 has suggests that are only used by a few geoms, so they are not listed in Imports, but only in 
    suggests, so that only in those specific scenarios, they will be required.   
3. Used only in vignettes or examples.   



####Devtools use_package()

```{r, eval=FALSE, results="hide"}
# use_package() will modify the DESCRIPTION
# and remind you how to use the function.

#Navigate to your new rv3 andopen .Rproj
#Probably a good idea to have a separate script in package to handle this.
devtools::use_package("ggplot2")
devtools::use_package("ggplot2", "suggests")

# load_all() will fail if imported packages are
# not installed. Up to you to ensure suggested
# packages have own checks.
# Installs all dependencies (including suggested)
devtools::install_deps(deps = TRUE)
```


NEVER use library() or require() in your package functions. This is the job of Imports in Description.  



###man/

Via roxygen2(). 

Note that roxygen comments have a special convention: regular comment with single quote: `#'`  

Roxygen turns R into Rd (R Documentation).  

General documentation workflow:  
1. Document  
2. devtools::document() or CTRL + SHIFT + D  
3. View  
4. Repeat 1:3 as required  
    * Note that `@export` updates can be handled in this manner as well; check NAMESPACE  


Order is always:  
1. Title (required)  
2. Description (optional)  
3. Details (optional)  

Most functions will also use:  
1. @param; describe inputs  
2. @examples; how it works  
3. @seealso; pointers to related functions  
4. @return; what it returns  
5. @export  

Look at RDocumentation.org  

####Formatting commands:

```{r, eval=FALSE, results="hide"}

\code{} Inline R code
\eqn{} Inline equation
(standard latex)
\emph{} Italic text
\strong{} Bold text

#' A bulleted list:
#' \itemize{
#' \item First item
#' \item Second item
#' }
#' An ordered list:
#' \enumerate{
#' \item First item
#' \item Second item
#' }

\link{foo} Link to foo in current package
\link[bar]{foo} Link to foo in package bar
\url{http://rstudio.com} Link to website
\href{http://rstudio.com}{Rstudio} Link to website with custom text
\email{hadley@rstudio.com} Email address
```


####To Preview links, workflow2

CTRL + SHIFT + B

Consider:  

```{r, eval=FALSE}
#In documentation for P (rv2): 
#` @seealso
#` \link{rv}
#
# save
# document()
# CTRL + SHIFT + L
# ?P #document updated, but link not active
# CTRL + SHIFT + B
# ?P #link now active

```


###Multiple functions in one file

For example, see `?rnorm`  
Useful for functions that are closely related

*Look into this some more*  
Check moments.R R for how this is done.  
Use `@name` for the first function  
Use `@rdname` for the next function(s), in the same script  
* Make sure that the names are the same throughout. 
* Use the name for the (set of) functions


Also, test `devtools::check_man()`

###Vignettes

Easiest way:  

```{r, eval= FALSE, results="hide"}
devtools::use_vignette("name")
# Adds to DESCRIPTION
Suggests: knitr
VignetteBuilder: knitr
# Creates vignettes/
# Drafts vignettes/name.Rmd
```


###Debugging

1. print()
2. debugonce()
3. browser()
4. options(error = recover)
5. options(warn = 2) #turn all warnings into errors


```{r, eval=FALSE}
f <- function(x) g(x)
g <- function(y) h(y)
h <- function(z) {
  stop("oops")
}

```


The goal is to find and fix problems.   

###Testing (automated test)

Why?  
1. check correctness
2. faster and more reliable than manual testing
3. stop bugs from coming back
    * so that one bugfix does not create another. 
4. reduce maintenance burden
5. enable refactoring
6. force good design
7. think adversarially
8. guard against future developers
9, communicate with colleagues
10. guard against complex interactions
11. social pressure
12. enable integration

Code that can be tested easily, and often is a consequence of modular design.  
When you stop working, leave a test failing-- makes it easier to pick up.  
Easier to make big changes without fear of accidentally breaking anything.  

####Unit testing with testthat

`length(revdep("testthat"))` #1395 packages as of today.  

So far, we have been doing exploratory coding.  
Automated testing using CTRL + SHIFT + T (`devtools::test()`)  

Write R Code, reload, test.  
Actually, you just write R code, test. No need to reload. Quite nice.  

Using test_me  

```{r, eval=FALSE}
library(cover)
#Runs all tests in your package and tells you what lines have not been tested. 
#Does not mean correct or not, just whether tested. 
#Probably overkill to aim for 100% test coverage-- focus on the really critical pieces. 
#This is how you get the coverage percentage.
#And get codecov

#How would you test the "P" function in rv2?
context("probability")
test_that("probability of coin flip is correct", {
  coin <- rv(1:2)
  
  expect_equal(P(coin == 1), 0.5)
  expect_equal(P(coin <=2), 1)
  expect_equal(P(coin < 0), 0)
})

#Shortcut to create test structure
devtools::create("./colSummary")
devtools::use_testthat()
devtools::use_test("test_colSummary")
use_package("purrr") #shortcut to add Imports to namespace with package(s)
#Insert roxygen skeleton works only if you have the fun, put cursor at beginning of fun, insert skeleton

#Regression tests: Not asserting what it should be, but what it used to be
expect_output_file() #spits out a file, and in the future, will tell you if the previous version is different from latest
test_that("rvs sum correctly", {
  dice <- rv(1:6)
  expect_output_file(print(dice + dice + dice + dice + dice),
  "sum_of_five_dice.txt", update = TRUE)
})

#Would test plot.rv() using vdiffr (upcoming package?)
```

Writing tests is the most difficult part.  

###NAMESPACE

Name conflicts-- if a name conflicts in two packages, the one called last wins by default; use double colon to be explicit.  

The other benefit of being explicit is that it insulates functions in your package from anything that might and likely will 
occur in users' namespace.  

A namespace splits functions into two classes: internal vs. external.  
By default, roxygen2 will export everything that does not belong with a dot.  
It is better to be explicit.  

Exporting the minimum number of functions is best practice, as it minimizes chances of namespace conflicts.  

Use the `@export` tag in roxygen2.  

Don't export internal helpers. 


###Imports

Particularly important for something like `magrittr %>%`.  
This is where you use `@import` in your actual function.  
The benefit of this is less typing. But it is less obvious when reading the code where these functions 
are coming from. 

Could also create an imports.R file to only import certain functions from certain packages. 
`#' @importFrom magrittr %>%` for example.  


```{r, eval=FALSE}
#' @import foo
#' @omport bar
fun <- function(x) {
  fun1(x) + fun2(x)
}

#Assume that fun1 and fun2 are from your package, and this works today. 
#Next year, the bar package adds fun1. Now you have a namespace conflict. 
#Generally, you should avoid using @import in your actual function scripts. 

#Things listed in NAMESPACE are basically functions. 

```

###R CMD CHECK

Useful even for local packages, even though false positives are possible.  
```{r, eval=FALSE}
#CTRL + SHIFT + E
#or
#devtools::check()
```


